### 2014-01-20 inst/doc/Advanced_writing_of_EML.Rmd
update tutorial

A similar situation arises with describing the coverage metadata element. 
```{r}
coverage <- eml_coverage(
  scientific_names = "Sarracenia purpurea", 
  dates            = c('2012-06-01', '2013-12-31'),
  geographic_description = "Harvard Forest Greenhouse, 
                            Tom Swamp Tract (Harvard Forest)", 
  NSEWbox          = c( 42.55,  42.42, -72.1, -72.29, 160, 330))
```

@@

```{r}
library(RWordXML)
library(XML)
f2 <- wordDoc(system.file("examples", "methods.docx", package="reml"))
doc <- f2[[getDocument(f2)]]
txt <- xpathSApply(doc, "//w:t", xmlValue)
## FIXME add <title> <section> and <para> blocking back: 
method_description <- paste(txt, collapse = "\n\n") 
```

@@

```{r}
methods <- new("methods")
methods@methodsStep@description <- method_description
```

@@

We now have all of the elements we wished to define for the `dataset` metadata. We can assemble them into a `datset` node as follows:

```{r}
dataset <- new("dataset", 
                title = title,
                creator = creator,
                contact = contact,
                pubDate = pubDate,
                associatedParty = other_researchers,
                keywordSet = keys,
                coverage = coverage,
                methods = methods,
                dataTable = c(dataTable))

@@

```{r}
eml <- new("eml", 
            dataset = dataset,
            additionalMetadata = additionalMetadata)
```

This call automatically adds a few additional bits of information, such as a unique identifier `packageId` for the metadata file.  This `eml` object we have just created now has the complete R representation of the EML data.  This is the same class of object returned by `eml_read`, and consequently we can use all of the read methods provided to explore or manipulate it.

@@

When we used the `eml_write()` function for the first time in the [Basic tutorial](), we simply passed a `data.set` object.  Internally the function made the remaining calls to `new("dataTable")`, `new("dataset")`, and `new("eml")` for us.  This time, the `eml_write` function can recognize that this is already an `eml` object, which it immediately converts to the XML representation and writes out as a file.  (If no file name is provided, the XML text will be returned to the console).

>>>>

A similar situation arises with describing the coverage metadata element. While we could call `new("coverage")` and build the element up slot by slot, most users will find the helper function more convenient:

```{r}
coverage <- eml_coverage(
  scientific_names = "Sarracenia purpurea", 
  dates            = c('2012-06-01', '2013-12-31'),
  geographic_description = "Harvard Forest Greenhouse, 
                            Tom Swamp Tract (Harvard Forest)", 
  NSEWbox          = c( 42.55,  42.42, -72.1, -72.29, 160, 330))
```

@@

```{r}
library(RWordXML)
library(XML)
f2 <- wordDoc(system.file("examples", "methods.docx", package="reml"))
doc <- f2[[getDocument(f2)]]
txt <- xpathSApply(doc, "//w:t", xmlValue)
## FIXME add <title> <section> and <para> blocking back: 
method <- paste(txt, collapse = "\n\n") 
```

@@

```{r}
methods <- new("methods", methodStep = c(new("methodStep", description = method)))
```

@@

We now have all of the elements we wished to define for the `dataset` metadata. A `dataset` can include yet more elements than we show here (see a list with `slotNames(new("dataset"))`, or see `?dataset` or the [EML documentation](http://knb.ecoinformatics.org/software/eml/eml-2.1.1/eml-dataset.html#DatasetType) for details). We can assemble them into a `dataset` node as follows:

```{r}
dataset <- new("dataset", 
                title = title,
                creator = creator,
                contact = contact,
                pubDate = pubDate,
                intellectualRights = rights,
                abstract = abstract,
                associatedParty = other_researchers,
                keywordSet = keys,
                coverage = coverage,
                methods = methods,
                dataTable = c(dataTable))

@@

```{r}
eml <- new("eml",
            packageId = uuid::UUIDgenerate(),
            system = "uuid", # type of identifier
            dataset = dataset,
            additionalMetadata = additionalMetadata)
```

This `eml` object we have just created now has the complete R representation of the EML data.  This is the same class of object returned by `eml_read`, and consequently we can use all of the read methods provided to explore or manipulate it.

## `eml` helper function

The helper function `eml` provides a shortcut to this step-by-step construction while retaining most of the expressiveness.

```{r}
eml     <- eml( dataset = dat,
                title = title,
                creator = creator,
                contact = contact,
                pubDate = pubDate,
                associatedParty = other_researchers,
                intellectualRights = rights,
                abstract = abstract,
                keywordSet = keys,
                coverage = coverage,
                methods = method,
                additionalMetadata = additionalMetadata
              )
```

The helper function handles certain steps automatically, such as the creation of a `uuid` for the package identifier.  The required fields `creator` and `contact` will use values set by `eml_config` or can take text strings or R `person` objects as arguments, as well as eml's `creator` or `contact` classes.  In this case we pass a `dataTable` (permitting additional customization we used above such as the description field).  Alternatively this could be the more basic `data.set` object as the first argument (`dataset = dat`) or an already constructed EML `dataset` class.

Because package identifer, creator and contact fields are required for valid EML, this constructor function is often more convenient then the stepwise process shown above. Typically a single call to this function can provide all the metadata needed to properly annotate a `data.set`.    This is similar to the behavior of the `eml_write` function, but returns the R eml object rather than writing out to an XML file.  Additional fields can then be added or manipulated using the standard R S4 subsetting methods (see Advanced Parsing tutorial).

@@

When we used the `eml_write()` function for the first time in the [Basic tutorial](), we simply passed a `data.set` object.  Internally the function made the remaining calls to `new("dataTable")`, `new("dataset")`, and `eml()` for us.  This time, the `eml_write` function can recognize that this is already an `eml` object, which it immediately converts to the XML representation and writes out as a file.  (If no file name is provided, the XML text will be returned to the console).
